\section{Static Analysis}

\subsection{Control Flow Graph}
\begin{itemize}
    \item Every node is an instruction
    \item If instruction $p$ is followed by $q$, there is an arc $p\rarr q$
\end{itemize}

True and false values in conditionals are not represented.
\begin{itemize}
    \item Assignment and reading is replaced with \emph{define}
    \item Variable in the right part is replaced with \emph{use}
\end{itemize}

Every instruction $p$ has $dep(p)$ and $use(p)$ set.

\subsection{Liveness intervals of Variables}
A variable is live at some point if later its current value is used by an instruction.

For the final node (if it has only one exiting arc): $live_{out}(p) = \emptyset$. For the other nodes:
\begin{align*}
    live_{in}(p) &= use(p) \cup (live_{out}(p) \setminus def(p)) \\
    live_{out}(p) &= \bigcup_{q\in succ(p)} live_{in}(q)
\end{align*}

To solve the system of flow equations, start with $live_{out}(p) = \emptyset$, then iteratively find $live_{in}(p)$ and $live_{out}(p)$, until convergence.

\paragraph{Applications}
\begin{itemize}
    \item Memory allocation: if two variables are never simultaneously live, they do not interfere
    \item Useless definitions: instruction that defines a variable that is not live out of the instruction.
\end{itemize}

\subsection{Reaching Definitions}

A definition of $a$ in $q$ reaches the entrance of instruction $p$ if there exists a path from $q$ to $p$ that does not traverse any node where $a$ is defined.

If $p$ defines $a$, any other definition $a_q$ (with $q\ne p$) is suppressed by $p$: $sup(p) = \{a_q | a \in def(p) \land a \in def(q) \land q \ne q\}$

\begin{align*}
    in(1) &= \emptyset \\
    out(p) &= def(p) \cup (in(p) \setminus sup(p)) \\
    in(p) &= \bigcup_{q\in pred(p)} out(q)
\end{align*}

Note that $in(p)$ and $out(p)$ contain $a_q$ or $a_?$ (for usage of variable $a$ at the start of the program).

To solve the system of flow equations, start with $in(p) = \emptyset$, then iteratively find $out(p)$ and $in(p)$, until convergence.

Constant propagation: it's possible to replace variable $a$ in $p$ with $k$ if it exists $q: a = k$ such that $a_q$ reaches $p$ and no other definition $a_r$ ($r\ne q$) reaches the entrance of $p$.

\subsection{Availability}

A variable $a$ is available on the entrance of $p$ if every path from the initial node to $p$ includes a definition of $a$.
